Different forms of uncertainty differently affect the evolution of social learning

### 2022-10-11 ms.tex
Updates from Overleaf

We also tracked and analyzed average net payoffs of agent populations across
all uncertainty parameters, and compared the simulated payoffs to 
all-social learner or all-asocial learner populations. 
Figure~\ref{fig:payoffs} shows the average net payoffs normalzied by lifespan
for our simulations and two reference conditions where agents are initialized
as homogenous populations fixated either as all social or all asocial learners.
Often the simulated payoffs follow the payoffs from the better-performing
homogenous group, but there are several important deviations to understand.
First, we observed frequency dependence of social learning, where 
agents from our main simulations who evolved to be social learners 
outperform agents in the homogenous social learner populations. This is most
pronounced when $L=1$, across all other uncertainty parameters, and even more
pronounced when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}, middle right).
In these cases, the presence of individual learners before social learning fixation
enabled social learners in our simulations to outperform all-social learner populations.
Next, we observed cases where long-lived populations ($L=8$) in simple environments
($B=2$) evolved to be individual learners, even though all-social learner populations
outperformed all-individual learner populations (Figure~\ref{fig:payoffs}, upper left).
By inspecting time series of individual model runs, we observe that
asocial learning fixates in response to a period of protracted environmental
instability, which causes social learning to more frequently provide
outdated information (Figure~\ref{fig:timeseries}a).
Conversely, sometimes when lifespan was short ($L=1$) and environmental variability
was moderately high ($u\approx0.5$), a period of environmental
stability led populations to fixate to social learning, despite individual learning
being the superior long-term strategy (Figure~\ref{fig:payoffs}, center top and middle). 
This behavior was again confirmed by examining time series of simultations
initialized with the commensurate uncertainty parameters, which show that
social learning fixates during periods of protratcted environmental stability
(Figure~\ref{fig:timeseries}b).

@@

Our model helps to demonstrate that social learning can act as a scaffold for asocial learning---despite the framing of many other papers, they are not opposite strategies. 
We endowed agents with a biologically-plausible individual learning mechanism, softmax search. This allowed for highly successful asocial learning that helped agents recover from the potential cost of receiving outdated information. This in turn enabled the evolution of social learning 
even when social information was likely to be outdated. \cm{this seems to say the opposite of the topic sentence. In topic sentence social learning-> asocial, but the mechanism descibed later is asocial learning -> social. Both seem possible, though the latter is easier for me to understand.}
Detailed theoretical models like this are likely to be critical for understanding how humans might adapt to an uncertain and rapidly changing world.
Such understanding could suggest plausibly beneficial interventions to mitigate existential threats~\citep{Moya2020,Jones2021}, especially in
the most vulnerable communities~\citep{McNamara2020}, and to capitalize on new behavioral opportunities such as transitioning to clean energy use and
production~\citep{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}. \cm{should we draw out some implications? I find this difficult, but e.g. stressing environmental change may discourage learning from older gen? having more tech options available for dealing with climate change may encourage social learning? Frankly so many of these problems seem to come back to encouraging people to socially learn from the RIGHT people, not just socially or asocially learning.}

>>>>

We also tracked and analyzed average net payoffs of agent populations across
all uncertainty parameters, and compared the simulated payoffs to 
homogenous all-social learner or all-asocial learner populations. Homogenous payoffs
were calculated by initializing simulation populations to be all social or all asocial learners,
then running the model for 100 generations so payoffs could stabilize; we then averaged 
the net payoffs from the final generation across 1000 trials to calculate the average
homogenous payoffs.

Figure~\ref{fig:payoffs} shows the average net payoffs normalized by lifespan
for our simulations and two reference conditions where agents are initialized
as homogenous populations fixated either as all social or all asocial learners.
Often the simulated payoffs follow the payoffs from the better-performing
homogenous group, but there are several important deviations to understand.
First, we observed frequency dependence of social learning, where 
agents from our main simulations who evolved to be social learners 
outperform agents in the homogenous social learner populations. This is most
pronounced when $L=1$, across all other uncertainty parameters, and even more
pronounced when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}, middle right).
In these cases, the presence of individual learners before social learning fixation
enabled social learners in our simulations to outperform all-social learner populations.
Next, we observed cases where long-lived populations ($L=8$) in simple environments
($B=2$) evolved to be individual learners, even though all-social learner populations
outperformed all-individual learner populations (Figure~\ref{fig:payoffs}, upper left).
By inspecting time series of individual model runs, we observe that
asocial learning fixates in response to a period of protracted environmental
instability, which causes social learning to more frequently provide
outdated information (Figure~\ref{fig:timeseries}a).
Conversely, sometimes when lifespan was short ($L=1$) and environmental variability
was moderately high ($u\approx0.5$), a period of environmental
stability led populations to fixate to social learning, despite individual learning
being the superior long-term strategy (Figure~\ref{fig:payoffs}, center top and middle). 
This behavior was again confirmed by examining time series of simultations
initialized with the commensurate uncertainty parameters, which show that
social learning fixates during periods of protratcted environmental stability
(Figure~\ref{fig:timeseries}b).

@@

Our model helps to demonstrate that social learning can act as a scaffold for individual learning---they are not opposite strategies. 
We endowed agents with a biologically-plausible individual learning mechanism, softmax search. This enabled agents to recover from the potential cost of receiving outdated information. This in turn enabled the evolution of social learning 
even when social information was likely to be outdated. 
% \cm{this seems to say the opposite of the topic sentence. In topic sentence social learning-> asocial, but the mechanism descibed later is asocial learning -> social. Both seem possible, though the latter is easier for me to understand.}
Detailed theoretical models like this are likely to be critical for understanding how humans might adapt to an uncertain and rapidly changing world.
Such understanding could suggest plausibly beneficial interventions to mitigate existential threats~\citep{Moya2020,Jones2021}, especially in
the most vulnerable communities~\citep{McNamara2020}, and to capitalize on new behavioral opportunities such as transitioning to clean energy use and
production~\citep{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}. \cm{should we draw out some implications? I find this difficult, but e.g. stressing environmental change may discourage learning from older gen? having more tech options available for dealing with climate change may encourage social learning? Frankly so many of these problems seem to come back to encouraging people to socially learn from the RIGHT people, not just socially or asocially learning.}
